---
title: "oculina16s_communitycomp_ancom"
author: "Ana Dulskiy"
date: "8/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ANCOM

Tutorial [here](https://github.com/FrederickHuangLin/ANCOM)

## Setup

```{r ancom packages, eval=FALSE}
library(readr)
library(tidyverse)
#library(dplyr)
library(nlme)
#install.packages('compositions')
library(compositions)
library(nloptr)
#ancom isn't an R package so you need to go to the ancom github and download the ancom.R file and save into wd
source("ancom.R")
```


```{r ancom all, eval=FALSE}
# shouldn't use rarefied data according to authors
# remove rows from ps.cleanest for unrarerified data

load("~/oculina/data/ps.clean.rev.Rdata")

#subset coral
ps.cleanest.new.coral <- subset_samples(ps.cleanest.new,type=="coral")
otu_data_unt <- data.frame(ps.cleanest.new.coral@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.cleanest.new.coral@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE 
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#setwd("~/oculina/community_comp/ancom")
#saveRDS(res.ss, file="ancom.res.ss.nonnest.RDS")

# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig  #0 W statistic for symb state?


#using new categorizations makes no difference for symbstate analysis
```

```{r ancom coral season}

#season
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.all = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.all, file="ancom.res.season.RDS")
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.all$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.all$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #8 taxa above 0.7
```

```{r ancom coral site}

#site
main_var = "site"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.site  = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.site, file="ancom.res.site.RDS")
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.site$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.site$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig 
#fig is comparing NR and RI to CL; two taxa for each above 0.7
```


## ANCOM by site{.tabset}

Note: shouldn't be done on rarefied according to authors

Ran these three chunks once, then loading in data below

### NR


```{r ancom NR symb state, eval=FALSE}
ps.less.coral.NR <- subset_samples(ps.less.coral,site=="NR")
otu_data_unt <- data.frame(ps.less.coral.NR@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.NR@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.NR,file="ancom.res.ss.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #only one value above 0? weird
```


```{r ancom NR season, eval=FALSE}
###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.NR,file="ancom.res.season.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #a few above zero
```

### RI

```{r ancom RI symb state, eval=FALSE}
ps.less.coral.RI <- subset_samples(ps.less.coral,site=="RI")
otu_data_unt <- data.frame(ps.less.coral.RI@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.RI@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.RI,file="ancom.res.ss.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero
```

```{r ancom RI season, eval=FALSE}
###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.RI,file="ancom.res.season.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #some sig
```

```{r ancom CL symb state, eval=FALSE}
ps.less.coral.CL <- subset_samples(ps.less.coral,site=="CL")
otu_data_unt <- data.frame(ps.less.coral.CL@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.CL@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.CL = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.RI,file="ancom.res.ss.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero
```

### CL
```{r ancom CL, eval=FALSE}
ps.clean.coral.CL <- subset_samples(ps.clean.coral,site=="CL")
otu_data_unt <- data.frame(ps.clean.coral.CL@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.clean.coral.CL@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.CL = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.CL,file="ancom.res.ss.CL.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.CL$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.CL$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero
```

## Synthesizing ANCOM results

Re-read in data

```{r load ancom results}
res.ss <- readRDS("./ancom/ancom.res.ss.nonnest.RDS")
res.season <- readRDS("./ancom/ancom.res.season.RDS")
res.site <- readRDS("./ancom/ancom.res.site.RDS")
res.ss.NR <- readRDS("./ancom/ancom.res.ss.NR.RDS")
res.ss.RI <- readRDS("./ancom/ancom.res.ss.RI.RDS")
res.season.NR <- readRDS("./ancom/ancom.res.season.NR.RDS")
res.season.RI <- readRDS("./ancom/ancom.res.season.RI.RDS")
```
```

Which ones are 'significant'

```{r subset out files}
#symb state
ss.out <- res.ss$out
ss.out.sig <- ss.out[ss.out$detected_0.6==TRUE,]
ss.out.sig #0
NR.ss.out <- res.ss.NR$out
NR.ss.out.sig <- NR.ss.out[NR.ss.out$detected_0.6==TRUE,] #0
RI.ss.out <- res.ss.RI$out
RI.ss.out.sig <- RI.ss.out[RI.ss.out$detected_0.6==TRUE,] #0

#season
season.out <- res.season$out
season.out.sig <- season.out[season.out$detected_0.6==TRUE,]
season.out.sig #9
NR.season.out <- res.season.NR$out
NR.season.out.sig <- NR.season.out[NR.season.out$detected_0.6==TRUE,] #8
RI.season.out <- res.season.RI$out
RI.season.out.sig <- RI.season.out[RI.season.out$detected_0.6==TRUE,] #4

#site
site.out <- res.site$out
site.out.sig <- site.out[site.out$detected_0.6==TRUE,]
site.out.sig #3

```
