---
title: "oculina16s_communitycomp"
output: html_document
date: '2022-06-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup

```{r packages composition}
library(rlang)
library(stringr)
library(dplyr)
library(stats)
library(ggpubr)
library(vegan)
library(cowplot)
library(tidyverse)
#library(MCMC.OTU)
#install.packages("remotes")
#remotes::install_github("Jtrachsel/funfuns")
library("funfuns")
```


Read in data

```{r read in data}
setwd("~/oculina/community_comp")
samdf <- read.csv("oculina16s_sampledata_plusneg_types.csv",header=TRUE)
load("taxa2.Rdata")
#load("ps.clean.Rdata")
load("ps.rare.Rdata")
#load("ps.rare.trim.Rdata")
#load("ps.trim.Rdata")
#ps.cleanest <- readRDS("phyloseq.cleanest.RDS")
```


Rename ASVs to be more informative

```{r rename ASVs}
tax <- as.data.frame(ps.rare@tax_table@.Data)
tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "D_0__",""),
                        Phylum = str_replace(tax[,2], "D_1__",""),
                        Class = str_replace(tax[,3], "D_2__",""),
                        Order = str_replace(tax[,4], "D_3__",""),
                        Family = str_replace(tax[,5], "D_4__",""),
                        Genus = str_replace(tax[,6], "D_5__",""),
                        Species = str_replace(tax[,7], "D_6__",""),
                        stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""


for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
####### Fill holes in the tax table
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}

tax_table(ps.rare) <- as.matrix(tax.clean)

```


# Core vs. accessory

## Core

```{r packages core, echo=FALSE}
#BiocManager::install("microbiome")
#remotes::install_github("r-lib/rlang")
library(microbiome)
```

```{r subset coral all}
#***I DON'T THINK I NEED THIS CODE HERE
ps.coral <- subset_samples(ps.rare,type=="coral")
#ps.coral.NR <- subset_samples(ps.coral,site=="NR")
#ps.coral.RI <- subset_samples(ps.coral,site=="RI")
#ps.coral.CL <- subset_samples(ps.coral,site=="CL")
#ps1.coral.site <- merge_samples(ps.coral, "site")
#ps2.coral.site <- transform_sample_counts(ps1.coral.site, function(x) x / sum(x))
#ps1.coral.ss <- merge_samples(ps.coral, "symbstate")
#ps2.coral.ss <- transform_sample_counts(ps1.coral.ss, function(x) x / sum(x))
#ps1.coral.seas <- merge_samples(ps.coral, "season")
#ps2.coral.seas <- transform_sample_counts(ps1.coral.seas, function(x) x / sum(x))
```


```{r create core ps}
pseq.core <- core(ps.rare, detection = 0, prevalence = .7)
pseq.core #4 taxa and 61 samples for ps.rare.trim, 5 taxa and 70 samples for ps.trim, 5 taxa and 64 samples for ps.rare
#saving pseq.core for ps.rare.trim
#setwd("~/oculina/community_comp")
#core.tax <- data.frame(pseq.core@tax_table)
#write.csv(core.tax,"core.taxa.trim.csv")

#grouping by all_id (just for coral for now)
ps.coral <- subset_samples(ps.rare,type=="coral") #need for accessory analysis later
ps.core.coral <- subset_samples(pseq.core,type=="coral")
ps_glom.coral <- tax_glom(ps.core.coral, "Genus")
ps0.coral <- transform_sample_counts(ps_glom.coral, function(x) x / sum(x))

#ps core coral grouped by all_id
ps1.coral <- merge_samples(ps0.coral, "all_id")
ps.core.coral.allid <- transform_sample_counts(ps1.coral, function(x) x / sum(x))

#ps grouped by site (because all_id shows every site-symb-season combination)
ps1.core.coral.site <- merge_samples(ps.core.coral, "site")
ps.core.coral.site <- transform_sample_counts(ps1.core.coral.site, function(x) x / sum(x))

#ps grouped by season
ps1.core.coral.season <- merge_samples(ps.core.coral, "season")
ps.core.coral.season <- transform_sample_counts(ps1.core.coral.season, function(x) x / sum(x))

#ps grouped by symb state
ps1.core.coral.symbstate <- merge_samples(ps.core.coral, "symbstate")
ps.core.coral.symbstate <- transform_sample_counts(ps1.core.coral.symbstate, function(x) x / sum(x))

#by site (need these for when just plotting data from one site)
ps.core.NR <- subset_samples(pseq.core,site=="NR")
ps.core.coral.NR <- subset_samples(ps.core.coral,site=="NR")
ps.core.RI <- subset_samples(pseq.core,site=="RI")
ps.core.coral.RI <- subset_samples(ps.core.coral,site=="RI")
ps.core.CL <- subset_samples(pseq.core,site=="CL")
ps.core.coral.CL <- subset_samples(ps.core.coral,site=="CL")

#ps relative abundance core coral
ps.core.coral.rel <- transform_sample_counts(ps.core.coral.allid, function(x) x / sum(x))

```

```{r bar plot core coral by all_id, site, season, & symbstate}

# by all_id (not relative)
gg.bar.coral.core.genus <- plot_bar(ps.core.coral.allid, fill="Genus")+
  geom_bar(stat="identity")+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))+
  scale_fill_brewer(palette="BrBG")

#site
gg.bar.coral.core.site.genus <- plot_bar(ps.core.coral.site, fill="Genus")+
  geom_bar(stat="identity")+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="BrBG")

gg.bar.coral.core.season.genus <- plot_bar(ps.core.coral.season, fill="Genus")+
  geom_bar(stat="identity")+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="BrBG")

gg.bar.coral.core.symbstate.genus <- plot_bar(ps.core.coral.symbstate, fill="Genus")+
  geom_bar(stat="identity")+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="BrBG")

gg.bar.coral.core.genus
gg.bar.coral.core.site.genus
gg.bar.coral.core.season.genus
gg.bar.coral.core.symbstate.genus

#ggsave(gg.bar.coral.core.genus,file="./figures/gg.bar.coral.core.genus.pdf",height=4)
#ggsave(gg.bar.coral.core.site.genus,file="./figures/gg.bar.coral.core.site.genus.pdf",height=4)
#ggsave(gg.bar.coral.core.season.genus,file="./figures/gg.bar.coral.core.season.genus.pdf",height=4)
#ggsave(gg.bar.coral.core.symbstate.genus,file="./figures/gg.bar.coral.core.symbstate.genus.pdf",height=4)

```

```{r PCoA plots coral core abundance}
# by all_id (NOT USEFUL, DON'T SAVE)
#pcoa.coral.core <- 
#  plot_ordination(ps.core.coral, ordinate(ps.core.coral,"PCoA", "bray"),color="all_id", shape="all_id")+
#  stat_ellipse()+
#  theme_cowplot()+
#  scale_color_brewer(palette="BrBG")+
#  xlab("Axis 1 (40.5%)")+
#  ylab("Axis 2 (19%)")+
#  ggtitle("Rarefied (Coral)")

# by site
pcoa.coral.core.site <- plot_ordination(ps.core.coral, ordinate(ps.core.coral,"PCoA", "bray"),color="site", shape="site")+
  stat_ellipse()+
  theme_cowplot()+
  scale_color_manual(name="Site",values=c("darkslategray3","darkslategray4","#000004"))+
  scale_shape_manual(name="Site",values=c(8,4,9))+
  xlab("Axis 1 (40.5%)")+
  ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")

pcoa.coral.core.season <- plot_ordination(ps.core.coral, ordinate(ps.core.coral,"PCoA", "bray"),color="season", shape="season")+
  stat_ellipse()+
  theme_cowplot()+
  scale_color_manual(name="Season",values=c("darkorchid4","goldenrod1"))+
  scale_shape_manual(name="Season",values=c(8,4))+
  xlab("Axis 1 (40.5%)")+
  ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")

pcoa.coral.core.symbstate <- plot_ordination(ps.core.coral, ordinate(ps.core.coral,"PCoA", "bray"),color="symbstate", shape="symbstate")+
  stat_ellipse()+
  theme_cowplot()+
  scale_colour_manual(name="Symbiotic State",values=c("lightcyan3","coral4"),labels=c("Aposymbiotic","Symbiotic")) +
  scale_shape_manual(name="Symbiotic State",values=c(8,4),labels=c("Aposymbiotic","Symbiotic"))+
 # xlab("Axis 1 (40.5%)")+
 # ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")

pcoa.coral.core.site
pcoa.coral.core.season
pcoa.coral.core.symbstate

#ggsave(pcoa.coral.core.site,file="./figures/pcoa.coral.core.site.pdf",height=4)
#ggsave(pcoa.coral.core.season,file="./figures/pcoa.coral.core.season.pdf",height=4)
#ggsave(pcoa.coral.core.symbstate,file="./figures/pcoa.coral.core.symbstate.pdf",height=4)

```
```{r PCoA coral core all sites}
#season
pcoa.coral.core.season.all <- plot_ordination(ps.core.coral,ordinate(ps.core.coral,"PCoA", "bray"),color="season", shape="season")+
  stat_ellipse()+
  theme_cowplot()+
  scale_colour_manual(name="Season", values=c("#8405A7FF","#ED7953FF"))+
  scale_shape_manual(name="Season",values=c(8,4))+
  xlab("Axis 1 (40.5%)")+
  ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))

#symbstate
pcoa.coral.core.symbstate.all <- plot_ordination(ps.core.coral,ordinate(ps.core.coral,"PCoA", "bray"),color="symbstate",shape="symbstate")+
  stat_ellipse()+
  theme_cowplot()+
  scale_colour_manual(name="Symbiotic State",values=c("lightcyan3","coral4"),labels=c("Aposymbiotic","Symbiotic")) +
  scale_shape_manual(name="Symbiotic State",values=c(8,4),labels=c("Aposymbiotic","Symbiotic"))+
  xlab("Axis 1 (40.5%)")+
  ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))

pcoa.coral.core.season.all
pcoa.coral.core.symbstate.all

#ggsave(pcoa.coral.core.season.all,file="./figures/pcoa.coral.core.season.all.pdf",width=8,height=4)
#ggsave(pcoa.coral.core.symbstate.all,file="./figures/pcoa.coral.core.symbstate.all.pdf",width=9,height=5)

```

```{r PCoA plots core all types}
pcoa.core.alltypes <- plot_ordination(pseq.core,ordinate(pseq.core,"PCoA", "bray"),color="type", shape="type")+
  stat_ellipse()+
  theme_cowplot()+
  scale_color_manual(name="Type",values=c("bisque2","paleturquoise3","chocolate4"))+
  scale_shape_manual(name="Type",values=c(8,4,9))+
  xlab("Axis 1 (25.8%)")+
  ylab("Axis 2 (22.5%)")+
  ggtitle("Rarefied (Coral)")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))

pcoa.core.alltypes
 
#ggsave(pcoa.core.alltypes,file="./figures/pcoa.core.alltypes.pdf",width=8,height=4)

 
```


```{r calculate core abundance, eval=FALSE}
# I DON'T THINK I NEED THIS CODE
# calculating core abundances #
core.sqs <- tax_table(pseq.core)
core.sqs.ids <- row.names(core.sqs)
core.sqs.ids
ps.rare.trim.rel <- transform_sample_counts(ps.rare.trim, function(x) x / sum(x))
seq.rare.rel <- data.frame(otu_table(ps.rare.trim.rel))
tax.core <- tax_table(ps.rare.trim.rel)
seq.core <- seq.rare.rel %>% select(c(core.sqs.ids))
core.rel <- data.frame(colMeans(seq.core))
total.rel <- data.frame(colMeans(seq.rare.rel))
total.rel.ordered <- data.frame(total.rel[order(-total.rel$colMeans.seq.rare.rel.),,drop=FALSE])
seq.core <- data.frame(otu_table(pseq.core))
seq.core <- data.frame(otu_table(pseq.core.rel))
dist.core <- vegdist(seq.core)
samdf.core <- data.frame(sample_data(pseq.core))
row.names(samdf.core)==row.names(seq.core)
bet.all <- betadisper(dist.core,samdf.core$zone)
```

### Core stats
```{r calculate coral core abundance, eval=FALSE}
# calculating core abundances #
core.coral.sqs <- tax_table(ps.core.coral)
core.coral.sqs.ids <- row.names(core.coral.sqs)
core.coral.sqs.ids
ps.coral.rare.rel <- transform_sample_counts(ps.coral, function(x) x / sum(x))
seq.coral.rare.rel <- data.frame(otu_table(ps.coral.rare.rel))
tax.core.coral <- tax_table(ps.coral.rare.rel)
seq.core.coral <- seq.coral.rare.rel %>% select(c(core.coral.sqs.ids))
core.coral.rel <- data.frame(colMeans(seq.core.coral))
total.coral.rel <- data.frame(colMeans(seq.coral.rare.rel))
total.coral.rel.ordered <- data.frame(total.rel[order(-total.coral.rel$colMeans.seq.coral.rare.rel.),,drop=FALSE])
```

```{r core stats}
seq.core.coral <- data.frame(otu_table(ps.core.coral))
#seq.core.coral <- data.frame(otu_table(ps.core.coral.rel)) *when I did this, row.names were all false below -- this didn't happen when I initially ran this code for ps.rare (all sample types), so should double check this later*
dist.core.coral <- vegdist(seq.core.coral)
samdf.core.coral <- data.frame(sample_data(ps.core.coral))
row.names(samdf.core.coral)==row.names(seq.core.coral)
bet.all <- betadisper(dist.core.coral,samdf.core.coral$site)
#anova(bet.all)
plot(bet.all) #mostly overlap
permutest(bet.all, pairwise = TRUE, permutations = 999) #not sig for any
# Pairwise comparisons:
# (Observed p-value below diagonal, permuted p-value above diagonal)
#      CL      NR    RI
#CL         0.94000 0.504
#NR 0.92569         0.630
#RI 0.50830 0.60444       

bet.all <- betadisper(dist.core.coral,samdf.core.coral$season)
plot(bet.all) #really overlap
permutest(bet.all, pairwise = TRUE, permutations = 999) #not sig for any
#          Fall Summer
#Fall             0.39
#Summer 0.36725   

adonis(seq.core.coral ~ site, data=samdf.core.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)  
#site       2    1.0470 0.52349  2.3414 0.10971  0.026 *
#Residuals 38    8.4960 0.22358         0.89029         
#Total     40    9.5429                 1.00000 

adonis(seq.core.coral ~ season, data=samdf.core.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)
#season     1    0.3211 0.32115  1.3582 0.03365  0.229
#Residuals 39    9.2218 0.23646         0.96635       
#Total     40    9.5429                 1.00000   

adonis(seq.core.coral ~ symbstate, data=samdf.core.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model    R2 Pr(>F)
#symbstate  1    0.3340 0.33404  1.4146 0.035  0.217
#Residuals 39    9.2089 0.23613         0.965       
#Total     40    9.5429                 1.000   

adonis(seq.core.coral ~ season*site, data=samdf.core.coral, permutations=999)
#            Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)  
#season       1    0.3211 0.32115  1.4549 0.03365  0.210  
#site         2    0.9782 0.48910  2.2158 0.10251  0.032 *
#season:site  1    0.2971 0.29708  1.3458 0.03113  0.256  
#Residuals   36    7.9465 0.22074         0.83271         
#Total       40    9.5429                 1.00000   

pairwise.adonis(seq.core.coral, factors = samdf.core.coral$site, permutations = 999)
#  pairs   F.Model     R2       p.value  p.adjusted
#NR vs RI	2.489729	0.07663126   	0.036	   0.036  *
#NR vs CL	3.080753	0.13347715  	0.021	   0.021  *
#RI vs CL	1.617150	0.05855601  	0.158    0.158

```

### Core stats by site

```{r core stats site}
#subset by site (done above)
#could try rel abundance too:
  #ps.core.coral.NR <- subset_samples(ps.core.coral.rel,site=="NR")
  #ps.core.coral.RI <- subset_samples(ps.core.coral.rel,site=="RI")
  #ps.core.coral.CL <- subset_samples(ps.core.coral.rel,site=="CL")

#### NR ####
seq.core.coral.NR <- data.frame(otu_table(ps.core.coral.NR))
dist.core.coral.NR <- vegdist(seq.core.coral.NR)
samdf.core.coral.NR <- data.frame(sample_data(ps.core.coral.NR))
row.names(samdf.core.coral.NR)==row.names(seq.core.coral.NR)

#season
bet.all <- betadisper(dist.core.coral.NR,samdf.core.coral.NR$season)
anova(bet.all) #ns
plot(bet.all) #overlap, not sig
adonis(seq.core.coral.NR ~ season,data=samdf.core.coral.NR, permutations=999) #not sig

#symb state
bet.all <- betadisper(dist.core.coral.NR,samdf.core.coral.NR$symbstate)
anova(bet.all)
plot(bet.all) #partially overlap but look quite different
adonis(seq.core.coral.NR ~ symbstate,data=samdf.core.coral.NR, permutations=999) #not sig

#### RI ####
seq.core.coral.RI <- data.frame(otu_table(ps.core.coral.RI))
dist.core.coral.RI <- vegdist(seq.core.coral.RI)
samdf.core.coral.RI <- data.frame(sample_data(ps.core.coral.RI))
row.names(samdf.core.coral.RI)==row.names(seq.core.coral.RI)

#season
bet.all <- betadisper(dist.core.coral.RI,samdf.core.coral.RI$season)
anova(bet.all) #ns
plot(bet.all) #some overlap
adonis(seq.core.coral.RI ~ season,data=samdf.core.coral.RI, permutations=999) #not sig

#symb state
bet.all <- betadisper(dist.core.coral.RI,samdf.core.coral.RI$symbstate)
anova(bet.all) #not sig
plot(bet.all) #overlap quite a bit
adonis(seq.core.coral.RI ~ symbstate,data=samdf.core.coral.RI, permutations=999) #PR = 0.1, so almost kinda sig?

#### CL ####
seq.core.coral.CL <- data.frame(otu_table(ps.core.coral.CL))
dist.core.coral.CL <- vegdist(seq.core.coral.CL)
samdf.core.coral.CL <- data.frame(sample_data(ps.core.coral.CL))
row.names(samdf.core.coral.CL)==row.names(seq.core.coral.CL)

#symb state
bet.all <- betadisper(dist.core.coral.CL,samdf.core.coral.CL$symbstate)
anova(bet.all) #not sig
plot(bet.all) #don't overlap that much
adonis(seq.core.coral.CL ~ symbstate,data=samdf.core.coral.CL, permutations=999) #not sig

```

## Accessory

```{r accessory}
ps.coral.otu <- data.frame(ps.coral@otu_table)
#ps.rare.trim.otu <- data.frame(ps.rare.trim@otu_table)
core.coral.tax <- data.frame(ps.core.coral@tax_table)
core.coral.ids <- c(rownames(core.coral.tax))
ps.coral.acc.otu <- ps.coral.otu[,!colnames(ps.coral.otu) %in% core.coral.ids ]
row.names(samdf) <- samdf$id
#remake phyloseq object
ps.acc.coral <- phyloseq(otu_table(ps.coral.acc.otu, taxa_are_rows=FALSE), 
                         sample_data(samdf), 
                         tax_table(taxa2))
ps.acc.coral #4809 taxa accessory
pcoa.acc.coral.all.site <- plot_ordination(ps.acc.coral,ordinate(ps.acc.coral,"PCoA", "bray"),color="site", shape="site")+
  stat_ellipse()+
  theme_cowplot()+
  scale_color_manual(name="Site",values=c("darkslategray3","darkslategray4","#000004"))+
  scale_shape_manual(name="Site",values=c(8,4,9))+
  xlab("Axis 1 (10.2%)")+
  ylab("Axis 2 (7.7%)")+
  ggtitle("Rarefied (Coral)")
#ggsave(pcoa.acc.coral.all.site,file="./figures/pcoa.acc.coral.all.site.pdf",height=4)

plot_ordination(ps.acc.coral,ordinate(ps.acc.coral,"PCoA", "bray"),color="symbstate", shape="symbstate")+
  stat_ellipse()+
  theme_cowplot()

plot_ordination(ps.acc.coral,ordinate(ps.acc.coral,"PCoA", "bray"),color="season", shape="season")+
  stat_ellipse()+
  theme_cowplot()
```

```{r stats acc - site}
#seq.acc <- data.frame(otu_table(ps.acc))
ps.acc.coral.rel <- transform_sample_counts(ps.acc.coral, function(x) x / sum(x))
seq.acc.coral <- data.frame(otu_table(ps.acc.coral.rel))
dist.acc.coral <- vegdist(seq.acc.coral)
samdf.acc.coral <- data.frame(sample_data(ps.acc.coral))
row.names(samdf.acc.coral)==row.names(seq.acc.coral)
bet.all <- betadisper(dist.acc.coral,samdf.acc.coral$site)
anova(bet.all) #sig*
plot(bet.all) 
permutest(bet.all, pairwise = TRUE, permutations = 999) #RI sig diff from NR and CL
  #Pairwise comparisons:
  #(Observed p-value below diagonal, permuted p-value above
  #diagonal)
  #      CL         NR      RI
  #CL           0.9890000 0.006
  #NR 0.9922739           0.031
  #RI 0.0072632 0.0300554      

bet.all <- betadisper(dist.acc.coral,samdf.acc.coral$season)
#anova(bet.all) #not sig
permutest(bet.all, pairwise = TRUE, permutations = 999) #no diff in disp
plot(bet.all) 

bet.all <- betadisper(dist.acc.coral,samdf.acc.coral$symbstate)
anova(bet.all) #not sig
permutest(bet.all, pairwise = TRUE, permutations = 999) 
plot(bet.all) 

adonis(seq.acc.coral ~ site, data=samdf.acc.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
#site       2     1.477 0.73848  1.9123 0.09145  0.001 ***
#Residuals 38    14.674 0.38616         0.90855           
#Total     40    16.151                 1.00000    
adonis(seq.acc.coral ~ site/season, data=samdf.acc.coral, permutations=999)
#            Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
#site         2     1.477 0.73848  2.0178 0.09145  0.001 ***
#site:season  2     1.499 0.74949  2.0479 0.09281  0.001 ***
#Residuals   36    13.175 0.36598         0.81574           
#Total       40    16.151                 1.00000           

adonis(seq.acc.coral ~ season, strata=samdf.acc.coral$site,data=samdf.acc.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
#season     1    1.1153 1.11530  2.8929 0.06905  0.001 ***
#Residuals 39   15.0357 0.38553         0.93095           
#Total     40   16.1510                 1.00000     

adonis(seq.acc.coral ~ symbstate, strata=samdf.acc.coral$site,data=samdf.acc.coral, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)
#symbstate  1     0.450 0.44996  1.1177 0.02786  0.137
#Residuals 39    15.701 0.40259         0.97214       
#Total     40    16.151                 1.00000  

pairwise.adonis(seq.acc.coral, factors = samdf.acc.coral$site, permutations = 999) #p < 0.001***
# pairs     F.Model    R2      p.value p.adjusted
#NR vs RI	1.858796	0.05834482	0.001  	0.001
#NR vs CL	2.029699	0.09213467	0.001  	0.001
#RI vs CL	1.891417	0.06781360	0.002  	0.002

```

```{r accessory by site and symbstate}
#by site
ps.acc.coral.NR <- subset_samples(ps.acc.coral,site=="NR")
ps.acc.coral.RI <- subset_samples(ps.acc.coral,site=="RI")
ps.acc.coral.CL <- subset_samples(ps.acc.coral,site=="CL")

pcoa.coral.acc.symbstate.all <- plot_ordination(ps.acc.coral,ordinate(ps.acc.coral,"PCoA", "bray"),color="symbstate", shape="symbstate")+
  stat_ellipse()+
  theme_cowplot()+
  scale_colour_manual(name="Symbiotic State",values=c("lightcyan3","coral4"),labels=c("Aposymbiotic","Symbiotic")) +
  scale_shape_manual(name="Symbiotic State",values=c(8,4))+
  xlab("Axis 1 (10.2%)")+
  ylab("Axis 2 (7.7%)")+
  ggtitle("Rarefied (Coral)")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
pcoa.coral.acc.symbstate.all

#ggsave(pcoa.coral.acc.symbstate.all,file="./figures/pcoa.coral.acc.symbstate.all.pdf",width=9,height=4)

```

```{r accessory by site and season}
pcoa.coral.acc.season.all <- plot_ordination(ps.acc.coral,ordinate(ps.acc.coral,"PCoA", "bray"),color="season", shape="season")+
  stat_ellipse()+
  theme_cowplot()+
  scale_colour_manual(name="Season", values=c("#8405A7FF","#ED7953FF"))+
  scale_shape_manual(name="Season",values=c(8,4))+
  #xlab("Axis 1 (40.5%)")+
  #ylab("Axis 2 (19%)")+
  ggtitle("Rarefied (Coral)")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))

pcoa.coral.acc.season.all
#ggsave(pcoa.coral.acc.season.all,file="./figures/pcoa.coral.acc.season.all.pdf",width=9,height=4)

```

```{r by site coral NR}
seq.acc.coral.NR <- data.frame(otu_table(ps.acc.coral.NR))
dist.acc.coral.NR <- vegdist(seq.acc.coral.NR)
samdf.acc.coral.NR <- data.frame(sample_data(ps.acc.coral.NR))
row.names(samdf.acc.coral.NR)==row.names(seq.acc.coral.NR)
#symbstate
bet.all <- betadisper(dist.acc.coral.NR,samdf.acc.coral.NR$symbstate)
anova(bet.all) #not sig
plot(bet.all) # mostly overlap
adonis(seq.acc.coral.NR ~ symbstate,data=samdf.acc.coral.NR, permutations=999) #not sig
#season
bet.all <- betadisper(dist.acc.coral.NR,samdf.acc.coral.NR$season)
anova(bet.all) #not sig
plot(bet.all) 
adonis(seq.acc.coral.NR ~ season,data=samdf.acc.coral.NR, permutations=999) #ns

```

```{r by site coral RI}
seq.acc.coral.RI <- data.frame(otu_table(ps.acc.coral.RI))
dist.acc.coral.RI <- vegdist(seq.acc.coral.RI)
samdf.acc.coral.RI <- data.frame(sample_data(ps.acc.coral.RI))
row.names(samdf.acc.coral.RI)==row.names(seq.acc.coral.RI)
#symbstate
bet.all <- betadisper(dist.acc.coral.RI,samdf.acc.coral.RI$symbstate)
anova(bet.all) #not sig
plot(bet.all) # some overlap
adonis(seq.acc.coral.NR ~ symbstate,data=samdf.acc.coral.NR, permutations=999) #not sig
#season
bet.all <- betadisper(dist.acc.coral.RI,samdf.acc.coral.RI$season)
anova(bet.all) 
plot(bet.all) #NO OVERLAP
adonis(seq.acc.coral.RI ~ season,data=samdf.acc.coral.RI, permutations=999)
#          Df SumsOfSqs MeanSqs F.Model     R2 Pr(>F)  
#season     1    0.9911 0.99111  2.6029 0.13278  0.001 ***
#Residuals 17    6.4731 0.38077         0.86722           
#Total     18    7.4642                 1.00000  

```

```{r by site coral CL}
seq.acc.coral.CL <- data.frame(otu_table(ps.acc.coral.CL))
dist.acc.coral.CL <- vegdist(seq.acc.coral.CL)
samdf.acc.coral.CL <- data.frame(sample_data(ps.acc.coral.CL))
row.names(samdf.acc.coral.CL)==row.names(seq.acc.coral.CL)
#symbstate
bet.all <- betadisper(dist.acc.coral.CL,samdf.acc.coral.CL$symbstate)
anova(bet.all) #not sig
plot(bet.all) # some overlap
adonis(seq.acc.coral.NR ~ symbstate,data=samdf.acc.coral.NR, permutations=999) #not sig
```


# Bar plots [needs fixing] {.tabset}

## All, by Phylum

```{r bar plot, eval=FALSE}
#Doesn't work great for my data -- not sure if it's supposed to look how it did or if it's just bc I have so many taxa
ps.allid <- merge_samples(ps.coral, "all_id")
ps.rel.allid <- transform_sample_counts(ps.allid, function(x) x / sum(x))

#all id
plot_bar(ps.rel.allid, fill="Phylum")+
  geom_bar(stat="identity")

#rel by site
plot_bar(ps2.coral.site, fill="Phylum")+
  geom_bar(stat="identity")

#rel by symbstate
plot_bar(ps2.coral.ss, fill="Phylum")+
  geom_bar(stat="identity")

#rel by season
plot_bar(ps2.coral.seas, fill="Phylum")+
  geom_bar(stat="identity")

ps.all.tab <- psmelt(ps.coral)%>%
  filter(!is.na(Abundance))%>%
  group_by(site,season,symbstate,all_id,Phylum,OTU)%>%
  summarize_at("Abundance",mean)

#ps.all.tab$site[ps.all.tab$site == "NR"] <- "NR"
#ps.all.tab$site[ps.all.tab$site == "RI"] <- "RI"
#ps.all.tab$site[ps.all.tab$site == "CL"] <- "CL"
#ps.all.tab$season[ps.all.tab$season == "Summer"] <- "Summer"
#ps.all.tab$season[ps.all.tab$season == "Fall"] <- "Fall"
gg.bar.all.phy <- ggplot(ps.all.tab,aes(x=site,y=Abundance,fill=Phylum))+
  geom_bar(stat="identity")+
  theme_cowplot()+
  #theme(legend.position="none")+
  facet_wrap(~site)
gg.bar.all.phy
#ggsave(gg.bar.all,file="bac.bar.all.pdf",height=4)
```


# ANCOM

Tutorial [here](https://github.com/FrederickHuangLin/ANCOM)

## Setup

```{r ancom packages}
library(readr)
library(tidyverse)
#library(dplyr)
library(nlme)
#install.packages('compositions')
library(compositions)
library(nloptr)
#ancom isn't an R package so you need to go to the ancom github and download the ancom.R file and save into wd
source("ancom.R")
```


```{r ancom all, eval=FALSE}
#ps.trim.coral <-  subset_samples(ps.less.trim,type=="coral")
ps.rare.trim.coral <-  subset_samples(ps.rare.trim,type=="coral")
#otu_data_unt <- data.frame(ps.trim.coral@otu_table)
otu_data_unt <- data.frame(ps.rare.trim.coral@otu_table)
otu_data<- data.frame(t(otu_data_unt))
#might need to make the sample names an actual row yet, not sure
meta_data = data.frame(ps.rare.trim.coral@sam_data)
#meta_data = data.frame(ps.clean.coral@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE 
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
#adj_formula = NULL; rand_formula = NULL
adj_formula = NULL; rand_formula = "~ 1 | site"
res.ss.all = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.all, file="ancom.res.ss.all.nonnest.RDS")

# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.all$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.all$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig  #0 W statistic for symb state?

#season
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = "~ 1 | site"
res.season.all = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.all, file="ancom.res.season.all.RDS")
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.all$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.all$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig  


```

## ANCOM by site{.tabset}

Note: shouldn't be done on rarefied according to authors

Ran these three chunks once, then loading in data below

### NR


```{r ancom NR, eval=FALSE}
ps.clean.coral.NR <- subset_samples(ps.clean.coral,site=="NR")
otu_data_unt <- data.frame(ps.clean.coral.NR@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.clean.coral.NR@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.NR,file="ancom.res.ss.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #only one value above 0? weird

```

```{r}
###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.NR,file="ancom.res.season.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig 
```



### RI
```{r ancom RI, eval=FALSE}
#load("ps.less.trim.Rdata") #just has those 2 useless samples removed
ps.trim.coral.RI <- subset_samples(ps.trim,site=="RI")
otu_data_unt <- data.frame(ps.trim.coral.RI@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.trim.coral.RI@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.RI,file="ancom.res.ss.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero

###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.RI,file="ancom.res.season.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #a few above 0.7
```

### CL
```{r ancom CL, eval=FALSE}
ps.clean.coral.CL <- subset_samples(ps.clean.coral,site=="CL")
otu_data_unt <- data.frame(ps.clean.coral.CL@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.clean.coral.CL@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.CL = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.CL,file="ancom.res.ss.CL.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.CL$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.CL$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero

```
## Synthesizing ANCOM results

Re-read in data

```{r load ancom results}
res.ss.NR <- readRDS("ancom.res.ss.NR.RDS")
res.ss.RI <- readRDS("ancom.res.ss.RI.RDS")
res.ss.CL <- readRDS("ancom.res.ss.CL.RDS")
res.season.NR <- readRDS("ancom.res.season.NR.RDS")
res.season.RI <- readRDS("ancom.res.season.RI.RDS")
```

Which ones are 'significant'

```{r subset out files}
#symb state
NR.ss.out <- res.ss.NR$out
NR.ss.out.sig <- NR.ss.out[NR.ss.out$detected_0.6==TRUE,] #0

RI.ss.out <- res.ss.RI$out
RI.ss.out.sig <- RI.ss.out[RI.ss.out$detected_0.6==TRUE,] #0

CL.ss.out <- res.ss.CL$out
CL.ss.out.sig <- CL.ss.out[CL.ss.out$detected_0.6==TRUE,] #0

#season
NR.season.out <- res.season.NR$out
NR.season.out.sig <- NR.season.out[NR.season.out$detected_0.6==TRUE,] #8

RI.season.out <- res.season.RI$out
RI.season.out.sig <- RI.season.out[RI.season.out$detected_0.6==TRUE,] #4

```
