---
title: "oculina16s_communitycomp"
output: html_document
date: '2022-06-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup

```{r packages composition}
library(rlang)
library(stringr)
library(dplyr)
library(stats)
library(ggpubr)
library(vegan)
library(cowplot)
library(tidyverse)
#library(MCMC.OTU)
#install.packages("remotes")
#remotes::install_github("Jtrachsel/funfuns")
library("funfuns")
#BiocManager::install("microbiome")
#remotes::install_github("r-lib/rlang")
library(microbiome)
```


Read in data

```{r read in data}
setwd("~/oculina/data")
samdf <- read.csv("oculina16s_sampledata_symdens.csv",header=TRUE)
load("taxa2.Rdata")
#load("ps.clean.Rdata")
load("ps.rare.Rdata")
#load("ps.rare.trim.Rdata")
#load("ps.trim.Rdata")
#ps.cleanest <- readRDS("phyloseq.cleanest.RDS")

```


Rename ASVs to be more informative

```{r rename ASVs}
tax <- as.data.frame(ps.rare@tax_table@.Data)
tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "D_0__",""),
                        Phylum = str_replace(tax[,2], "D_1__",""),
                        Class = str_replace(tax[,3], "D_2__",""),
                        Order = str_replace(tax[,4], "D_3__",""),
                        Family = str_replace(tax[,5], "D_4__",""),
                        Genus = str_replace(tax[,6], "D_5__",""),
                        Species = str_replace(tax[,7], "D_6__",""),
                        stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""


for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
####### Fill holes in the tax table
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}

tax_table(ps.rare) <- as.matrix(tax.clean)


```



### Stats

# Transform data using centered log ratio (makes data symmetric and linearly related)
```{r transform data using CLR}
#library(microbiome)
ps.rare.clr <- transform(ps.rare, 'clr')
#subset coral
ps.coral <- subset_samples(ps.rare.clr,type=="coral")

```



```{r stats for coral}
seq.coral <- data.frame(otu_table(ps.coral))
dist.coral <- vegdist(seq.coral)
samdf.coral <- data.frame(sample_data(ps.coral))
row.names(samdf.coral)==row.names(seq.coral)
bet.all <- betadisper(dist.coral,samdf.coral$site)
#anova(bet.all)
plot(bet.all) #mostly overlap
permutest(bet.all, pairwise = TRUE, permutations = 999) #not sig for any
#Response: Distances
#          Df   Sum Sq   Mean Sq      F N.Perm Pr(>F)  
#Groups     2 0.047477 0.0237384 3.2692    999  0.054 .
#Residuals 38 0.275928 0.0072613                       

#Pairwise comparisons:
#(Observed p-value below diagonal, permuted p-value above diagonal)
#          CL        NR    RI
#CL           0.4670000 0.006
#NR 0.4528432           0.109
#RI 0.0069592 0.0950158    

bet.all <- betadisper(dist.coral,samdf.coral$season)
plot(bet.all)
permutest(bet.all, pairwise = TRUE, permutations = 999) #not sig at all

bet.all <- betadisper(dist.coral,samdf.coral$symbstate)
plot(bet.all)
permutest(bet.all, pairwise = TRUE, permutations = 999) #not sig

# adonis
adonis2(seq.coral ~ site, data=samdf.coral, permutations=999)
#         Df SumOfSqs      R2     F Pr(>F)    
#site      2   1.4922 0.10055 2.124  0.001 ***
#Residual 38  13.3483 0.89945                 
#Total    40  14.8406 1.00000     

adonis2(seq.coral ~ season, data=samdf.coral, permutations=999) #*** ??
adonis2(seq.coral ~ symbstate, data=samdf.coral, permutations=999) #not sig


```

### Stats by site
```{r}
#subset by site (done above)
  #ps.coral.NR <- subset_samples(ps.coral,site=="NR")
  #ps.coral.RI <- subset_samples(ps.coral,site=="RI")
  #ps.coral.CL <- subset_samples(ps.coral,site=="CL")

#### NR ####
seq.coral.NR <- data.frame(otu_table(ps.coral.NR))
dist.coral.NR <- vegdist(seq.coral.NR)
samdf.coral.NR <- data.frame(sample_data(ps.coral.NR))
row.names(samdf.coral.NR)==row.names(seq.coral.NR)

#season
bet.all <- betadisper(dist.coral.NR,samdf.coral.NR$season)
anova(bet.all) #ns
plot(bet.all)
permutest(bet.all, pairwise = TRUE, permutations = 999) #ns
adonis2(seq.coral.NR ~ season,data=samdf.coral.NR, permutations=999) #ns 

#symb state
bet.all <- betadisper(dist.coral.NR,samdf.coral.NR$symbstate)
anova(bet.all)
plot(bet.all) #partially overlap but look quite different
permutest(bet.all, pairwise = TRUE, permutations = 999) #ns
adonis2(seq.coral.NR ~ symbstate,data=samdf.coral.NR, permutations=999) #not sig


### RI ###
seq.coral.RI <- data.frame(otu_table(ps.coral.RI))
dist.coral.RI <- vegdist(seq.coral.RI)
samdf.coral.RI <- data.frame(sample_data(ps.coral.RI))
row.names(samdf.coral.RI)==row.names(seq.coral.RI)

#season
bet.all <- betadisper(dist.coral.RI,samdf.coral.RI$season)
anova(bet.all) #ns
plot(bet.all) #no overlap
permutest(bet.all, pairwise = TRUE, permutations = 999) #ns
adonis2(seq.coral.RI ~ season,data=samdf.coral.RI, permutations=999) 
#         Df SumOfSqs      R2      F Pr(>F)   
#season    1   0.9091 0.13282 2.6038  0.002 **
#Residual 17   5.9357 0.86718                 
#Total    18   6.8448 1.00000       

#symb state
bet.all <- betadisper(dist.coral.RI,samdf.coral.RI$symbstate)
anova(bet.all)
plot(bet.all)
permutest(bet.all, pairwise = TRUE, permutations = 999) #ns
adonis2(seq.coral.RI ~ symbstate,data=samdf.coral.RI, permutations=999) #ns

### CL ###
seq.coral.CL <- data.frame(otu_table(ps.coral.CL))
dist.coral.CL <- vegdist(seq.coral.CL)
samdf.coral.CL <- data.frame(sample_data(ps.coral.CL))
row.names(samdf.coral.CL)==row.names(seq.coral.CL)

#symb state
bet.all <- betadisper(dist.coral.CL,samdf.coral.CL$symbstate)
anova(bet.all)
plot(bet.all)
permutest(bet.all, pairwise = TRUE, permutations = 999) #ns
adonis2(seq.coral.CL ~ symbstate,data=samdf.coral.CL, permutations=999) #ns

```




# ANCOM

Tutorial [here](https://github.com/FrederickHuangLin/ANCOM)

## Setup

```{r ancom packages, eval=FALSE}
library(readr)
library(tidyverse)
#library(dplyr)
library(nlme)
#install.packages('compositions')
library(compositions)
library(nloptr)
#ancom isn't an R package so you need to go to the ancom github and download the ancom.R file and save into wd
source("ancom.R")
```


```{r ancom all, eval=FALSE}
# shouldn't use rarefied data according to authors
# remove rows from ps.cleanest for unrarerified data
#load("~/oculina/data/ps.cleanest.Rdata")
#seqtab.cleanest <- read.csv("~/oculina/data/oculina16s_seqtab.cleanest.csv",row.names=1)
#row.names.remove <- c("D6","F3","H2","F7","H4","H7")
#samdf.cleanest <- data.frame(ps.cleanest@sam_data)
#samdf.cleanest <- samdf.cleanest[!(row.names(samdf.cleanest) %in% row.names.remove), ]
#ps.cleanest.less <- phyloseq(otu_table(seqtab.cleanest, taxa_are_rows=FALSE), 
 #                        sample_data(samdf.cleanest), 
  #                       tax_table(taxa2))
#save(ps.cleanest.less,file="ps.cleanest.less.Rdata")

#subset coral
#ps.less.coral <- subset_samples(ps.cleanest.less,type=="coral")

#load("~/oculina/data/ps.trim.RData")
load("~/oculina/community_comp/ps.cleanest.less.Rdata")
ps.less.coral <- subset_samples(ps.cleanest.less,type=="coral")
#ps.trim.coral <- subset_samples(ps.trim,type=="coral")
otu_data_unt <- data.frame(ps.less.coral@otu_table)
#otu_data_unt <- data.frame(ps.trim.coral@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral@sam_data)
#meta_data = data.frame(ps.trim.coral@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE 
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
#adj_formula = NULL; rand_formula = NULL
adj_formula = NULL; rand_formula = "~ 1 | site"
res.ss.all = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.all, file="ancom.res.ss.all.nonnest.RDS")

# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.all$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.all$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig  #0 W statistic for symb state?

#season
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = "~ 1 | site"
res.season.all = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.all, file="ancom.res.season.all.RDS")
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.all$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.all$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig  

```

## ANCOM by site{.tabset}

Note: shouldn't be done on rarefied according to authors

Ran these three chunks once, then loading in data below

### NR


```{r ancom NR symb state, eval=FALSE}
ps.less.coral.NR <- subset_samples(ps.less.coral,site=="NR")
otu_data_unt <- data.frame(ps.less.coral.NR@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.NR@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.NR,file="ancom.res.ss.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #only one value above 0? weird

```


```{r ancom NR season, eval=FALSE}
###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.NR = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.NR,file="ancom.res.season.NR.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.NR$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.NR$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #a few above zero

```

### RI

```{r ancom RI symb state, eval=FALSE}
ps.less.coral.RI <- subset_samples(ps.less.coral,site=="RI")
otu_data_unt <- data.frame(ps.less.coral.RI@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.RI@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.RI,file="ancom.res.ss.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero

```

```{r ancom RI season, eval=FALSE}
###season
# Step 2: ANCOM
main_var = "season"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.season.RI = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.season.RI,file="ancom.res.season.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.season.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.season.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #some sig

```

```{r ancom CL symb state, eval=FALSE}
ps.less.coral.CL <- subset_samples(ps.less.coral,site=="CL")
otu_data_unt <- data.frame(ps.less.coral.CL@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.less.coral.CL@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.CL = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.RI,file="ancom.res.ss.RI.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.RI$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.RI$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero
```

### CL
```{r ancom CL, eval=FALSE}
ps.clean.coral.CL <- subset_samples(ps.clean.coral,site=="CL")
otu_data_unt <- data.frame(ps.clean.coral.CL@otu_table)
otu_data<- data.frame(t(otu_data_unt))
meta_data = data.frame(ps.clean.coral.CL@sam_data)
meta_data = meta_data %>% rename(Sample.ID = id)
# Step 1: Data preprocessing
feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

###symb state
# Step 2: ANCOM
main_var = "symbstate"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
res.ss.CL = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
#saveRDS(res.ss.CL,file="ancom.res.ss.CL.RDS")
# Step 3: Volcano Plot
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")
# Annotation data
dat_ann = data.frame(x = min(res.ss.CL$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")
fig = res.ss.CL$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig #all zero

```
## Synthesizing ANCOM results

Re-read in data

```{r load ancom results}
res.ss.NR <- readRDS("./ancom/ancom.res.ss.NR.RDS")
res.ss.RI <- readRDS("./ancom/ancom.res.ss.RI.RDS")
res.season.NR <- readRDS("./ancom/ancom.res.season.NR.RDS")
res.season.RI <- readRDS("./ancom/ancom.res.season.RI.RDS")
```

Which ones are 'significant'

```{r subset out files}
#symb state
NR.ss.out <- res.ss.NR$out
NR.ss.out.sig <- NR.ss.out[NR.ss.out$detected_0.6==TRUE,] #0

RI.ss.out <- res.ss.RI$out
RI.ss.out.sig <- RI.ss.out[RI.ss.out$detected_0.6==TRUE,] #0

#CL.ss.out <- res.ss.CL$out
#CL.ss.out.sig <- CL.ss.out[CL.ss.out$detected_0.6==TRUE,] #0

#season
NR.season.out <- res.season.NR$out
NR.season.out.sig <- NR.season.out[NR.season.out$detected_0.6==TRUE,] #8

RI.season.out <- res.season.RI$out
RI.season.out.sig <- RI.season.out[RI.season.out$detected_0.6==TRUE,] #4

```



```{r bubble plot}
library("reshape2")
library("stringr")
library("ggplot2")
library("RColorBrewer")

#setwd("~/oculina_old/setup/tax")
OTU = read.table("ASVtable.txt", header=TRUE, sep="\t", row.names = 1)
RAtab <- sweep(OTU,1,rowSums(OTU),"/")
#write.csv(RAtab, "ASVtablePercent.csv") #uncomment this to write out your new ASV table
#row sums before
rowSums(OTU)
#check row sums after-- should be 1 for all samples
rowSums(RAtab)


```

