---
title: "oculina16s_analysis"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: '2022-06-01'
editor_options: 
  chunk_output_type: console
---
### Based on DADA2 Pipeline 1.16 Walkthrough & Nicola Kriefall's Moorea holobiont analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

## Packages
```{r packages}
library(ggplot2)
library(cowplot)
library(phyloseq)
library(car)
library(ggpubr)
library(vegan)
library(dada2)
```

setwd("~/oculina/data/")

## Read in data

```{r read in data}
samdf <- read.csv("~/oculina/data/oculina16s_sampledata_plusneg_types.csv",header=TRUE) #R is making me put the full path all of a sudden, not sure why
row.names(samdf) <- samdf$id
load("~/oculina/data/taxa2.Rdata")
```


## Phyloseq objects

```{r making phyloseq objects, eval=FALSE}
#setwd("~/oculina/diversity")
seqtab.rare <- read.csv("oculina16s_seqtab.cleanest.rare_2k copy",row.names=1)
ps.rare <- phyloseq(otu_table(seqtab.rare, taxa_are_rows=FALSE), 
                    sample_data(samdf), 
                    tax_table(taxa2))
ps.rare 
#save(ps.rare,file="ps.rare.Rdata")
#4814 taxa during revisions - raw & rarefied

#raw, unrarefied version:
seqtab.clean <- read.csv("oculina16s_seqtab.cleanest.csv",row.names=1)
ps.clean <- phyloseq(otu_table(seqtab.clean, taxa_are_rows=FALSE), 
                    sample_data(samdf), 
                    tax_table(taxa2))
ps.clean
#save(ps.clean,file="ps.clean.Rdata")
#6573 taxa just raw, not rarefied

```

##TRIMMED VERSIONS:
```{r}
seqtab.rare.trim <- read.csv("oculina16s_seqtab.trim.rare_2k.csv",row.names=1)
ps.rare.trim <- phyloseq(otu_table(seqtab.rare.trim, taxa_are_rows=FALSE), 
                    sample_data(samdf), 
                    tax_table(taxa2))
ps.rare.trim
#save(ps.rare.trim,file="ps.rare.trim.Rdata")
#796 taxa &  61 samples rarefied

#checking what happens if we read in the unrarefied version:
seqtab.trim <- read.csv("oculina16s_seqtab.cleanest.trim.csv",row.names=1)
ps.trim <- phyloseq(otu_table(seqtab.trim, taxa_are_rows=FALSE), 
                    sample_data(samdf), 
                    tax_table(taxa2))
ps.trim
#save(ps.trim,file="ps.trim.Rdata")
#796 taxa & 70 samples not rarefied
```


```{r read in ps objects}
setwd("~/oculina/data")
#load("ps.clean.Rdata")
load("ps.rare.Rdata")
#load("ps.rare.trim.Rdata")
#load("ps.trim.Rdata")
```

# Diversity

[Notes from phyloseq author](https://rdrr.io/bioc/phyloseq/man/estimate_richness.html)
Visualize alpha-diversity - Should be done on raw, untrimmed dataset

```{r generate div metrics}
df <- data.frame(estimate_richness(ps.rare, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.clean, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.trim, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.rare.trim, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
df$id <- rownames(df)
df.div <- merge(df,samdf,by="id") #add sample data
#shannon diversity divided by species richness
df.div$even <- df.div$Shannon/(log(df.div$Observed))
df.div.NR <- subset(df.div,site=="NR")
df.div.RI <- subset(df.div,site=="RI")
df.div.CL <- subset(df.div,site=="CL")
df.div.summer <- subset(df.div,season=="Summer")
df.div.fall <- subset(df.div,season=="Fall")
df.div.coral <- subset(df.div, type=="coral")
df.div.ss <- subset(df.div, type=="sediment" | type =="seawater")
df.div.coral.NR <- subset(df.div.coral,site=="NR")
df.div.coral.RI <- subset(df.div.coral,site=="RI")
df.div.coral.CL <- subset(df.div.coral,site=="CL")
df.div.coral.NR.RI <- subset(df.div.coral,site %in% c("NR","RI"))
```


## Alpha metrics {.tabset}
 
Looking at everything individually, more of a synthesis below

### Shannon

```{r shannon site}
gg.site.sha <- ggplot(df.div.coral,aes(x=site,y=Shannon,color=site))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("Shannon index")+
  xlab("Site")+
  theme_cowplot()+
  theme(legend.position="none")
gg.site.sha
```


```{r shannon coral}
#plotting coral season & symb state
gg.season.symb <- ggplot(df.div.coral, aes(x=season,y=Shannon,color=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=symbstate), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  xlab("Season")+
  ylab("Shannon index")+
  #scale_x_discrete(labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("lightcyan3","coral4")) +
  theme_cowplot()+
  guides(season = "none")+
  labs(color = "Symbiotic State")+
  #geom_jitter(alpha=0.5)+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
gg.season.symb

```
```{r shannon sedsw}
#plotting sed and sw
gg.sedsw <- ggplot(df.div.ss, aes(x=season,y=Shannon,color=season))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(shape=type), size = 3, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  xlab("Season")+
  ylab("Shannon index")+
  #scale_x_discrete(labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("#ED7953FF","#8405A7FF"))+
  theme_cowplot()+
  guides(color = "none")+
  labs(shape = "Type")+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))

gg.sedsw
```

```{r}
gg.shannon.summer <- ggplot(df.div.summer, aes(x=site,y=Shannon,color=type))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=type), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  scale_colour_manual(values=c("darksalmon","paleturquoise3","chocolate4")) +
  ylab("Shannon Index")+
  xlab("Site")+
  #labs(color = "Symbiotic State")+
  theme_cowplot() +
  scale_x_discrete(limits=c('NR','RI', 'CL')) +
  theme(legend.position="none")+
  ylim(min_value=0, max_value=6.5)+
  ggtitle("a) Summer")

gg.shannon.fall <- ggplot(df.div.fall, aes(x=site,y=Shannon,color=type))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=type), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  scale_colour_manual(name="Type",values=c("darksalmon","paleturquoise3","chocolate4"), labels=c("Coral","Seawater","Sediment")) +
  ylab("Shannon Index")+
  xlab("Site")+
  #labs(color = "Symbiotic State")+
  theme_cowplot() +
  scale_x_discrete(limits=c('NR','RI'))+
  ylim(min_value=0, max_value=6.5)+
  ggtitle("b) Fall")

plot_grid(gg.shannon.summer, gg.shannon.fall, axis="bt",align="hv")
```


```{r shannon symb state}
gg.ss.sha <- ggplot(df.div.coral,aes(x=site,y=Shannon,color=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=symbstate), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  scale_colour_manual(values=c("lightcyan3","coral4")) +
  ylab("Shannon Index")+
  xlab("Site")+
  labs(color = "Symbiotic State")+
  theme_cowplot() +
  scale_x_discrete(limits=c('NR','RI', 'CL'))
  #theme(legend.position="none")
gg.ss.sha
```

```{r shannon revised stats (glm)}
glm(df.div.coral$Shannon~site+season)

```


```{r shannon stats coral}
#Shannon
shapiro.test(df.div$Shannon) #ns - normal distribution
shapiro.test(df.div.coral$Shannon) #ns
leveneTest(df.div$Shannon~site/season,data=df.div) #ns (variances equal)
leveneTest(df.div.coral$Shannon~site/season,data=df.div.coral) #ns (variances equal)


# Ok so t-test is for comparing two groups, (one-way) ANOVA for > 2 with one independent variable
# Use two-way ANOVA for > 1 independent variable
# So when comparing fall vs summer, or sym vs apo for all groups together, I need to use t-test
# When comparing sites, use ANOVA
# When comparing coral vs. sed vs. sw, use ANOVA
# For two-way ANOVA, need to test for homogeneity of variance (Levene's) and independence of observations (Kruskal-Wallis) as well as normality (Shapiro); nice explainer here: https://www.scribbr.com/statistics/two-way-anova/

# For two-way ANOVA:
  # I have three independent variables (for coral):
    # 1) Site
    # 2) Season
    # 3) Symb state
  # Three options for two-way ANOVA:
    # 1) Without interaction (just additive)
    # 2) With interaction (x * y)
    # 3) With blocking variable (I don't believe I have one, but maybe site?)
  # Make models for all and then compare via aictab()

aov.seas <-aov(Shannon~season, data=df.div.coral)
aov.ss <- aov(Shannon~symbstate, data=df.div.coral)
### Two-way (additive):
aov.add.all <- aov(Shannon~site+season+symbstate, data=df.div.coral)
summary(aov.add.all)
#without site
aov.add.seas.ss <-  aov(Shannon~season+symbstate, data=df.div.coral)
summary(aov.add.seas.ss) #season sig (.)

### Two-way (interactive):
aov.int.all <- aov(Shannon~site*season*symbstate, data=df.div.coral)
summary(aov.int.all) # site, season sig (.), site:season and site:symbstate sig (*)
#without site
aov.int.seas.ss <- aov(Shannon~season*symbstate, data=df.div.coral)
summary(aov.int.seas.ss) #season sig (.)

### Two-way (w/ site as blocking variable)
aov.block <- aov(Shannon~season*symbstate+site, data=df.div.coral)
summary(aov.block) #season sig (*)

### Use AIC to compare models
#install.packages("AICcmodavg")
#library(AICcmodavg)

model.set <- list(aov.seas,aov.ss,aov.add.all,aov.add.seas.ss,aov.int.all,aov.int.seas.ss,aov.block)
model.names <- c("aov.seas", "aov.ss", "aov.add.all", "aov.add.seas.ss", "aov.int.all", "aov.int.seas.ss", "aov.block")
aictab(model.set, modnames = model.names)
#                 K   AICc Delta_AICc AICcWt Cum.Wt     LL
#aov.seas         3  98.92       0.00   0.55   0.55 -46.14
#aov.add.seas.ss  4 101.30       2.37   0.17   0.71 -46.09
#aov.int.seas.ss  5 102.73       3.80   0.08   0.79 -45.51
#aov.add.all      6 102.99       4.06   0.07   0.86 -44.26
#aov.ss           3 103.03       4.11   0.07   0.93 -48.19
#aov.block        7 104.51       5.58   0.03   0.97 -43.56
#aov.int.all     11 104.58       5.66   0.03   1.00 -36.74

### Conclusion: season is the best predictor of diversity, with additive model excluding site second best (Shannon~season+symbstate)
### However, this data includes CL, which might be skewing the results since it only has one season of data. Going to redo this analysis below excluding CL.

```

```{r anova without CL}
aov.seas <-aov(Shannon~season, data=df.div.coral.NR.RI)
summary(aov.seas) #ns
aov.ss <- aov(Shannon~symbstate, data=df.div.coral.NR.RI)
summary(aov.ss) #ns
### Two-way (additive):
aov.add.all <- aov(Shannon~site+season+symbstate, data=df.div.coral.NR.RI)
summary(aov.add.all) #symbstate sig (.)
#without site
aov.add.seas.ss <-  aov(Shannon~season+symbstate, data=df.div.coral.NR.RI)
summary(aov.add.seas.ss) #ns

### Two-way (interactive):
aov.int.all <- aov(Shannon~site*season*symbstate, data=df.div.coral.NR.RI)
summary(aov.int.all) #season (*), site:season (*), site:symbstate(**)
#without site
aov.int.seas.ss <- aov(Shannon~season*symbstate, data=df.div.coral.NR.RI)
summary(aov.int.seas.ss) #ns

### Two-way (w/ site as blocking variable)
aov.block <- aov(Shannon~season*symbstate+site, data=df.div.coral.NR.RI)
summary(aov.block) #ns

### Use AIC to compare models

model.set <- list(aov.seas,aov.ss,aov.add.all,aov.add.seas.ss,aov.int.all,aov.int.seas.ss,aov.block)
model.names <- c("aov.seas", "aov.ss", "aov.add.all", "aov.add.seas.ss", "aov.int.all", "aov.int.seas.ss", "aov.block")
aictab(model.set, modnames = model.names)

###Best model is aov.seas (confusing b/c the summary didn't show significance?), second best is interactive including site

```




```{r wilcox tests, eval=FALSE}
# I don't think I should be doing a wilcox test because that is for non-parametric data, and my data is normally distributed
#stats by season (coral):
wilcox.test(Shannon~season, data=df.div.coral) # < 0.05 *
summary(aov(Shannon~season,data=df.div.coral)) # < 0.05 *
wilcox.test(Shannon~season,data=df.div.coral.NR) #ns
summary(aov(Shannon~season,data=df.div.coral.NR)) #ns
wilcox.test(Shannon~season,data=df.div.coral.RI) # < 0.01 **
summary(aov(Shannon~season,data=df.div.coral.RI)) # < 0.01 **
#stats symb state
wilcox.test(Shannon~symbstate,data=df.div.coral) #ns
summary(aov(Shannon~symbstate,data=df.div.coral)) #ns
wilcox.test(Shannon~symbstate,data=df.div.coral.NR) # < 0.1 .
summary(aov(Shannon~symbstate,data=df.div.coral.NR)) # < 0.5 *
wilcox.test(Shannon~symbstate,data=df.div.coral.RI) #ns
summary(aov(Shannon~symbstate,data=df.div.coral.RI)) # < 0.1 .
wilcox.test(Shannon~symbstate,data=df.div.coral.CL) #ns
summary(aov(Shannon~symbstate,data=df.div.coral.CL)) #ns
```

```{r shannon no CL stats}
shapiro.test(df.div.coral.NR.RI$Shannon) #ns - data normal
leveneTest(df.div.coral.NR.RI$Shannon~site/season,data=df.div.coral.NR.RI) #ns (variances equal)
a.div <- aov(Shannon~site/season,data=df.div.coral.NR.RI)
summary(a.div) #site:season significant (<0.05)
a.div.symb <- aov(Shannon~symbstate/site,data=df.div.coral.NR.RI)
summary(a.div.symb) #symbstate:site significant (<0.05)
a.div.symb.seas <- aov(Shannon~symbstate/season,data=df.div.coral.NR.RI)
summary(a.div.symb.seas) #ns
TukeyHSD(a.div) # ns
TukeyHSD(a.div.symb) # sig <0.1 for symbiotic:RI-symbiotic:NR 
TukeyHSD(a.div.symb.seas) #ns

#stats by season (coral):
wilcox.test(Shannon~symbstate,data=df.div.coral.NR.RI) #ns
wilcox.test(Shannon~season,data=df.div.coral.NR.RI) # ns
summary(aov(Shannon~season,data=df.div.coral.NR.RI)) #ns
wilcox.test(Shannon~symbstate,data=df.div.coral.NR.RI) # ns
summary(aov(Shannon~symbstate,data=df.div.coral.NR.RI)) #ns

```


### Simpson

```{r simpson site}
gg.site.sim <- ggplot(df.div.coral,aes(x=site,y=InvSimpson,color=site))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("Inv. Simpson index")+
  xlab("Site")+
  theme_cowplot()+
  theme(legend.position="none")
gg.site.sim
```

```{r simpson season}
gg.season.sim <- ggplot(df.div.coral,aes(x=season,y=InvSimpson,color=season))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("Inv. Simpson index")+
  xlab("Site")+
  theme_cowplot()+
  theme(legend.position="none")
gg.season.sim
```


```{r simpson stats}
#simpson
shapiro.test(df.div.coral$InvSimpson) #not normal
df.div.coral$si.log <- log(df.div.coral$InvSimpson) 
shapiro.test(df.div.coral$si.log) #normal, ns

##season
leveneTest(df.div.coral$si.log~site/season,data=df.div.coral) #ns
a.div.season.coral <- aov(si.log~site/season,data=df.div.coral)
summary(a.div.season.coral) #site & site:season significant (<.05 *)
TukeyHSD(a.div.season.coral) #RI-summer:RI-Fall sig (<0.05 *)

##symbstate
leveneTest(df.div.coral$si.log~site/symbstate,data=df.div.coral) #ns
a.div.symb.coral <- aov(si.log~site/symbstate,data=df.div.coral)
summary(a.div.symb.coral) #site significant (<0.1 .)
TukeyHSD(a.div.symb.coral) #ns

#stats by site (season):
wilcox.test(InvSimpson~season,data=df.div.coral) # < 0.01 **
wilcox.test(InvSimpson~season,data=df.div.coral.NR.RI) # < 0.05 *
wilcox.test(InvSimpson~season,data=df.div.coral.NR) #ns
summary(aov(InvSimpson~season,data=df.div.coral.NR)) #ns
wilcox.test(InvSimpson~season,data=df.div.coral.RI) # < 0.05 *
summary(aov(InvSimpson~season,data=df.div.coral.RI)) # < 0.05 *

#stats by site (symb state):
wilcox.test(InvSimpson~symbstate,data=df.div.coral) # ns
wilcox.test(InvSimpson~symbstate,data=df.div.coral.NR) # < 0.01 **
summary(aov(InvSimpson~symbstate,data=df.div.coral.NR)) # < 0.05 *
wilcox.test(InvSimpson~symbstate,data=df.div.coral.RI) # ns
summary(aov(InvSimpson~symbstate,data=df.div.coral.RI)) # < 0.1 .
wilcox.test(InvSimpson~symbstate,data=df.div.coral.CL) # ns
summary(aov(InvSimpson~symbstate,data=df.div.coral.CL)) # ns
```


### Richness

```{r richness site}
gg.site.obs <- ggplot(df.div.coral,aes(x=site,y=Observed,color=site))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("ASV richness")+
  xlab("Site")+
  theme_cowplot()+
  theme(legend.position="none")
gg.site.obs
```
```{r richness season}
gg.season.obs <- ggplot(df.div.coral, aes(x=season,y=Observed,color=season,shape=season))+
  geom_boxplot(outlier.shape=NA)+
  xlab("Reef zone")+
  ylab("ASV richness")+
  theme_cowplot()+
  scale_shape_manual(values=c(16,15),labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("#ED7953FF","#8405A7FF"),labels=c("Fall","Summer"))+
  theme(legend.position="none")+
  geom_jitter(alpha=0.5)+
  facet_wrap(~site)+
  scale_x_discrete(labels=c("Fall","Summer"))
gg.season.obs
```
```{r richness symbstate}
gg.symb.obs <- ggplot(df.div.coral, aes(x=symbstate,y=Observed,color=symbstate,shape=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  xlab("Reef zone")+
  ylab("ASV richness")+
  theme_cowplot()+
  scale_colour_manual(values=c("lightcyan3","coral4"),labels=c("Apo","Symb")) +
  scale_shape_manual(values=c(16,15),labels=c("Apo","Symb"))+
  theme(legend.position="none")+
  geom_jitter(alpha=0.5)+
  facet_wrap(~site)+
  scale_x_discrete(labels=c("Apo","Symb"))
gg.symb.obs
```
```{r richness symbstate season}
gg.season.symb.obs <- ggplot(df.div.coral, aes(x=season,y=Observed,color=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=symbstate), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  xlab("Season")+
  ylab("ASV Richness")+
  #scale_x_discrete(labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("lightcyan3","coral4")) +
  theme_cowplot()+
  guides(season = "none")+
  labs(color = "Symbiotic State")+
  #geom_jitter(alpha=0.5)+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
gg.season.symb.obs
```


```{r richness stats}
#otu richness
shapiro.test(df.div.coral$Observed) # p-value = 1.059e-05 ???
df.div.coral$obs.log <- log(df.div.coral$Observed)
shapiro.test(df.div.coral$obs.log) #p-value = 0.1021
leveneTest(df.div.coral$obs.log~site/season,data=df.div.coral) 
a.div <- aov(obs.log~site/season,data=df.div.coral)
summary(a.div) #ns         
TukeyHSD(a.div) #ns

#stats by site (season):
wilcox.test(Observed~season,data=df.div.coral.NR) #ns
summary(aov(Observed~season,data=df.div.coral.NR)) #ns
wilcox.test(Observed~season,data=df.div.coral.RI) #ns
summary(aov(Observed~season,data=df.div.coral.RI)) #ns

#stats by site (symbstate):
wilcox.test(Observed~symbstate,data=df.div.coral.NR) #ns
summary(aov(Observed~symbstate,data=df.div.coral.NR)) #ns
wilcox.test(Observed~symbstate,data=df.div.coral.RI) # < 0.05 *
summary(aov(Observed~symbstate,data=df.div.coral.RI)) # < 0.05 *
wilcox.test(Observed~symbstate,data=df.div.coral.CL) #ns
summary(aov(Observed~symbstate,data=df.div.coral.CL)) #ns
```


### Evenness

```{r evenness site}
gg.site.eve <- ggplot(df.div.coral,aes(x=site,y=even,color=site))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("Evenness")+
  xlab("Site")+
  theme_cowplot()+
  theme(legend.position="none")
gg.site.eve
```

```{r evenness season}
gg.season.eve <- ggplot(df.div.coral, aes(x=season, y=even,color=season,shape=season))+
  geom_boxplot(outlier.shape=NA)+
  xlab("Season")+
  ylab("Evenness")+
  theme_cowplot()+
  scale_shape_manual(values=c(16,15),labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("#ED7953FF","#8405A7FF"),labels=c("Fall","Summer"))+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(legend.position="none")+
  geom_jitter(alpha=0.5)+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
  #scale_x_discrete(labels=c("BR","FR"))
gg.season.eve
```

```{r evenness symbstate}
gg.symb.eve <- ggplot(df.div.coral, aes(x=symbstate, y=even,color=symbstate,shape=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  xlab("Symbiotic State")+
  ylab("Evenness")+
  theme_cowplot()+
  scale_shape_manual(values=c(16,15),labels=c("Apo","Sym"))+
  scale_colour_manual(values=c("lightcyan3","coral4"),labels=c("Apo","Symb")) +
  theme(legend.position="none")+
  geom_jitter(alpha=0.5)+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
gg.symb.eve
```
```{r evenness symbstate season}
gg.season.symb.eve <- ggplot(df.div.coral, aes(x=season,y=even,color=symbstate))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(aes(color=symbstate), size = 2, alpha = 0.3, position=position_jitterdodge(jitter.width=0.2)) +
  xlab("Season")+
  ylab("Evenness")+
  #scale_x_discrete(labels=c("Fall","Summer"))+
  scale_colour_manual(values=c("lightcyan3","coral4")) +
  theme_cowplot()+
  guides(season = "none")+
  labs(color = "Symbiotic State")+
  #geom_jitter(alpha=0.5)+
  facet_wrap(~forcats::fct_relevel(site, "NR", "RI", "CL"))
gg.season.symb.eve
```



```{r evenness stats}
#evenness
shapiro.test(df.div.coral$even) # < 0.05 *
leveneTest(df.div.coral$even~site*season,data=df.div.coral) #ns
leveneTest(df.div.coral$even~site*symbstate,data=df.div.coral) #ns

# aov season
a.div <- aov(even~site*season,data=df.div.coral)
summary(a.div)
#           Df  Sum Sq Mean Sq F value  Pr(>F)   
#site         2 0.07688 0.03844   5.281 0.00975 **
#season       1 0.06191 0.06191   8.505 0.00606 **
#site:season  1 0.03761 0.03761   5.166 0.02910 * 
#Residuals   36 0.26206 0.00728 
TukeyHSD(a.div) #ns

#aov site*symbstate
a.div <- aov(even~symbstate,data=df.div.coral)
summary(a.div) #ns
a.div <- aov(even~site,data=df.div.coral)
summary(a.div) # < 0.05 *
a.div <- aov(even~site*symbstate,data=df.div.coral)
summary(a.div) #ns
TukeyHSD(a.div) # NR-CL and RI-CL significant (<0.05 *)

#stats by site (season):
wilcox.test(even~season,data=df.div.coral.NR.RI) # < 0.01 ** becomes sig once CL taken out
summary(aov(even~season,data=df.div.coral.NR.RI)) # < 0.01 **
wilcox.test(even~season,data=df.div.coral.NR) #ns
summary(aov(even~season,data=df.div.coral.NR)) #ns
wilcox.test(even~season,data=df.div.coral.RI) #ns
summary(aov(even~season,data=df.div.coral.RI)) # 0 *** wow - all being driven by RI, not NR, but that's probably bc very few samples from NR for fall

#stats by site (symbstate):
wilcox.test(even~symbstate,data=df.div.coral.NR) # < 0.01 ** sig effect of symb state evenness @ NR but not RI or CL!
summary(aov(even~symbstate,data=df.div.coral.NR)) # < 0.01 **
wilcox.test(even~symbstate,data=df.div.coral.RI) #ns
summary(aov(even~symbstate,data=df.div.coral.RI)) #ns
wilcox.test(even~symbstate,data=df.div.coral.CL) #ns (p = 1!)
summary(aov(even~symbstate,data=df.div.coral.CL)) #ns
```


## Phylogenetic diversity (Faith's D)

Tutorial from dada2 author [here](https://f1000research.com/articles/5-1492/v2)

```{r packages phylo d, eval=FALSE}
#install.packages('devtools')
#library(devtools)
#devtools::install_github('twbattaglia/btools')
#library(btools)
```


## Playing with microshades
```{r microshades setup}
#remotes::install_github("KarstensLab/microshades")
library("microshades")
#remotes::install_github("mikemc/speedyseq")
library("speedyseq")

#all
ps_glom <- tax_glom(ps.rare, "Genus")
ps0 <- transform_sample_counts(ps_glom, function(x) x / sum(x))
ps1 <- merge_samples(ps0, "fullname")
ps.all <- transform_sample_counts(ps1, function(x) x / sum(x))

# include other variables in ps.all (beyond just fullname)
seqtab.cleanest <- read.csv("oculina16s_seqtab.cleanest.csv",row.names=1)
ps.cleanest <- phyloseq(otu_table(seqtab.cleanest, taxa_are_rows=FALSE), 
                    sample_data(samdf), 
                    tax_table(taxa2))
ps.cleanest 
#save(ps.cleanest,file="ps.cleanest.Rdata")
samdf.cleanest <- data.frame(ps.cleanest@sam_data)

#for rarefied data, remove everything < 2000
row.names.remove <- c("D6","F3","H2","F7","H4","H7")
samdf.rare <- samdf.cleanest[!(row.names(samdf.cleanest) %in% row.names.remove), ]

#variable1 = as.character(get_variable(samdf.rare, "type"))
#sample_data(ps.all)$type <- mapply(paste0, variable1)
all(rownames(samdf.rare) %in% sample_names(ps.all))
samdf.rare = sample_data(samdf.rare)
merge_phyloseq(ps.all, samdf.rare)
head(sample_data(ps.all))

#overall by all_id
ps_glom.rare <- tax_glom(ps.rare, "Genus")
ps0.rare <- transform_sample_counts(ps_glom.rare, function(x) x / sum(x))
ps1.rare <- merge_samples(ps0.rare, "all_id")
ps.rare.allid <- transform_sample_counts(ps1.rare, function(x) x / sum(x))

# overall by type
ps1.rare.type <- merge_samples(ps0.rare, "type")
ps.rare.type <- transform_sample_counts(ps1.rare.type, function(x) x / sum(x))

# overall coral by all_id
ps.coral <- subset_samples(ps.rare,type=="coral")
ps_glom.coral <- tax_glom(ps.coral, "Genus")
ps0.coral <- transform_sample_counts(ps_glom.coral, function(x) x / sum(x))
ps1.coral <- merge_samples(ps0.coral, "all_id")
ps.coral.allid <- transform_sample_counts(ps1.coral, function(x) x / sum(x))

# overall coral by site
ps1.coral.site <- merge_samples(ps.coral, "site")
ps.coral.site <- transform_sample_counts(ps1.coral.site, function(x) x / sum(x))

# overall coral by season
ps1.coral.season <- merge_samples(ps.coral, "season")
ps.coral.season <- transform_sample_counts(ps1.coral.season, function(x) x / sum(x))

# overall coral by symbstate
ps1.coral.symbstate <- merge_samples(ps.coral, "symbstate")
ps.coral.symbstate <- transform_sample_counts(ps1.coral.symbstate, function(x) x / sum(x))

# Use microshades function prep_mdf to agglomerate, normalize, and melt the phyloseq object
mdf.prep.rare <- prep_mdf(ps.rare, subgroup_level="Genus")
mdf.prep <- prep_mdf(ps.all, subgroup_level="Genus")
mdf.prep.rare.allid <- prep_mdf(ps.rare.allid, subgroup_level = "Genus")
mdf.prep.rare.type <- prep_mdf(ps.rare.type, subgroup_level = "Family")
mdf.prep.allid <- prep_mdf(ps.coral.allid, subgroup_level = "Family")
mdf.prep.site <- prep_mdf(ps.coral.site, subgroup_level = "Family")
mdf.prep.season <- prep_mdf(ps.coral.season, subgroup_level = "Family")
mdf.prep.symbstate <- prep_mdf(ps.coral.symbstate, subgroup_level = "Family")

# Create a color object for the specified data
color.obj.rare <- create_color_dfs(mdf.prep.rare, group_level = "Phylum", subgroup_level = "Genus", cvd = TRUE)
color.obj <- create_color_dfs(mdf.prep, group_level = "Phylum", subgroup_level = "Genus", cvd = TRUE)
color.obj.rare.allid <- create_color_dfs(mdf.prep.rare.allid, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)
color.obj.rare.type <- create_color_dfs(mdf.prep.rare.type, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)
color.obj.allid <- create_color_dfs(mdf.prep, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)
color.obj.site <- create_color_dfs(mdf.prep.site, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)
color.obj.season <- create_color_dfs(mdf.prep.season, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)
color.obj.symbstate <- create_color_dfs(mdf.prep.symbstate, group_level = "Phylum", subgroup_level = "Family", cvd = TRUE)


# Extract
mdf.rare <- color.obj.rare$mdf
cdf.rare <- color.obj.rare$cdf
mdf <- color.obj$mdf
cdf <- color.obj$cdf
mdf.rare.allid <- color.obj.rare$mdf #contains sample data and abundance info
cdf.rare.allid <- color.obj.rare$cdf #contains color mapping info
mdf.rare.type <- color.obj.rare.type$mdf 
cdf.rare.type <- color.obj.rare.type$cdf 
mdf.coral.allid <- color.obj$mdf.allid
cdf.coral.allid <- color.obj$cdf.allid
mdf.coral.site <- color.obj.site$mdf
cdf.coral.site <- color.obj.site$cdf
mdf.coral.season <- color.obj.season$mdf
cdf.coral.season <- color.obj.season$cdf
mdf.coral.symbstate <- color.obj.symbstate$mdf
cdf.coral.symbstate <- color.obj.symbstate$cdf
```


```{r microshades initial plots}
plot1 <- plot_microshades(mdf, cdf, group_label = "Phylum Genus")
plot1.all <- plot_microshades(mdf.rare, cdf.rare, group_label = "Phylum Family")
plot1.all.type <- plot_microshades(mdf.rare.type, cdf.rare.type, group_label = "Phylum Family")
plot1.site <- plot_microshades(mdf.coral.site, cdf.coral.site, group_label = "Phylum Family")
plot1.season <- plot_microshades(mdf.coral.season, cdf.coral.season, group_label = "Phylum Family")
plot1.symbstate <- plot_microshades(mdf.coral.symbstate, cdf.coral.symbstate, group_label = "Phylum Family")

```

```{r all samples reordered by group}
# reorder_samples_by will change the order of samples based on an abundance of a specified subgroup taxonomy
new.sample.order <- reorder_samples_by(mdf, cdf, order = "Thiovulum", group_level = "Phylum", subgroup_level = "Genus", sink_abundant_groups = TRUE)

mdf.new.sample.order <-new.sample.order$mdf
cdf.new.sample.order <-new.sample.order$cdf

plot_2 <- plot_microshades(mdf.new.sample.order, cdf.new.sample.order, group_label = "Phylum Family")

plot_2 + scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.key.size = unit(0.2, "cm"), text=element_text(size=10)) +
  theme(axis.text.x = element_text(size= 6))
```

```{r all_id reordered by group}
# reorder_samples_by will change the order of samples based on an abundance of a specified subgroup taxonomy
new.sample.order.coral <- reorder_samples_by(mdf.rare, cdf.rare, order = "Cyanobium_PCC-6307", group_level = "Phylum", subgroup_level = "Genus", sink_abundant_groups = TRUE)

mdf.new.sample.order <-new.sample.order.coral$mdf
cdf.new.sample.order <-new.sample.order.coral$cdf

plot_2 <- plot_microshades(mdf.new.sample.order, cdf.new.sample.order, group_label = "Phylum Genus")

plot_2 + scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.key.size = unit(0.2, "cm"), text=element_text(size=10)) +
  theme(axis.text.x = element_text(size= 6))
```

```{r reorder samples by abundance of group levels only}
# reorder_samples_by will change the order of samples based on an abundance of a specified subgroup taxonomy
# The default subgroup_level is "Genus" 
new.group.order <- reorder_samples_by(mdf.rare, cdf.rare, group_level = "Phylum", subgroup_level = "Family", sink_abundant_groups = TRUE)

mdf.new.group.order <-new.group.order$mdf
cdf.new.group.order <-new.group.order$cdf

plot_3 <- plot_microshades(mdf.new.group.order, cdf.new.group.order, group_label = "Phylum Family")

plot_3 + scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.key.size = unit(0.2, "cm"), text=element_text(size=10)) +
  theme(axis.text.x = element_text(size= 6))
```
```{r}
# reorder_samples_by will change the order of samples based on Family an abundance of a group and reorder the Phylum to sink the most abundant groups
reordered <- reorder_samples_by(mdf.rare, cdf.rare, order = "Other", group_level = "Phylum", subgroup_level = "Family",sink_abundant_groups=TRUE)

mdf.reordered <-reordered$mdf
cdf.reordered <-reordered$cdf

plot_3 <- plot_microshades(mdf.reordered, cdf.reordered, group_label = "Phylum Family")

plot_3 + scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.key.size = unit(0.2, "cm"), text=element_text(size=10)) +
  theme(axis.text.x = element_text(size= 6))
```


```{r facet wrap}
# remove current Sample column (currently a repeat of id), and then rename fullname = Sample
#library(dplyr)
mdf.rare <- mdf.rare %>%
  select(-Sample) %>%
  rename(Sample = fullname) %>%
  rename(number = sample)

plot.rare <- plot_microshades(mdf.rare, cdf.rare, group_label = "Phylum Genus")

plot.rare + 
  theme(legend.key.size = unit(0.5, "cm"), text=element_text(size=10)) +
  theme(axis.text.x = element_text(size= 6)) +
  facet_grid(~type, scales = "free_x", space="free") +
  theme (strip.text.x = element_text(size = 10))
  
```




